---
title: "final"
author: "esther"
date: "11/17/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(ggplot2)
library(dplyr)
library(lubridate)
library(repr)

options(repr.plot.width=5, repr.plot.height=4)
```
The Metro Interstate Traffic volume dataset is from UCI Machine Learning Repostitory, data is taken from Minneapolis-St Paul, MN traffic volume for westbound I-94. There are nine variables and 48,204 observations of every hour from 2012 to 2018. 

Variables:
holiday - categorical indicating if there was a holiday
temperature - temperature in kelvin
rain_1h - rain in mm that occured in the hour
snow_1h snow in mm that occured in the hour
clouds_all - cloud cover percentage
weather_main - description of weather
weather_descrription - long description
date_time - hour of data collected
traffic_volume - numeric hourly average true range of traffic volume

```{r data}
#read in data
traffic <- read_csv("Metro_Interstate_Traffic_Volume.csv")
head(traffic)
dim(traffic)
```

```{r adddinggg stuffff cuz im bored}
unique(traffic$weather_main)
#change columns into new dataframe df
df <- traffic %>%
        mutate(
          temp = temp - 270.15,
          weather_main = factor(weather_main),
          holiday = ifelse(holiday == "None", 0, holiday) %>% as.numeric,
          holiday = ifelse(is.na(holiday), 1, 0) %>% as.numeric)
```
```{r}
#new columns
df$date <- as.Date(traffic$date_time, "%m/%d/%Y")
df$day <- weekdays(df$date)
df$hour <- strftime(traffic$date_time, format="%H", tz = "CST")
df$hour2 <- as.numeric(df$hour)
```
```{r}
df$year <- strftime(traffic$date_time, format="%Y")
```
```{r}
df$tod <- ifelse(df$hour2 >= 03 & df$hour2 <= 07 , "Early Morning",
            ifelse(df$hour2 > 07 & df$hour2 <= 11, "Morning",
            ifelse(df$hour2 > 11 & df$hour2 <= 17, "Afternoon", 
            ifelse(df$hour2 > 17 & df$hour2 <= 20, "Evening", "Night"))))
```
```{r https://stackoverflow.com/a/46846474 source}
#mode of factor variables
calculate_mode <- function(x) {
  uniqx <- unique(x)
  uniqx[which.max(tabulate(match(x, uniqx)))]
}
```
```{r https://stackoverflow.com/a/9501225 source}
getSeason <- function(DATES) {
    WS <- as.Date("2012-12-15", format = "%Y-%m-%d") # Winter Solstice
    SE <- as.Date("2012-3-15",  format = "%Y-%m-%d") # Spring Equinox
    SS <- as.Date("2012-6-15",  format = "%Y-%m-%d") # Summer Solstice
    FE <- as.Date("2012-9-15",  format = "%Y-%m-%d") # Fall Equinox

    # Convert dates from any year to 2012 dates
    d <- as.Date(strftime(DATES, format="2012-%m-%d"))

    ifelse (d >= WS | d < SE, "Winter",
      ifelse (d >= SE & d < SS, "Spring",
        ifelse (d >= SS & d < FE, "Summer", "Fall")))
}
```
```{r}
df$season <- getSeason(df$date_time)
```
```{r}
df %>% mutate(tod = as.factor(tod))
```
```{r}
df$tod <- ordered(df$tod, levels = c("Early Morning", "Morning", "Afternoon", "Evening", "Night"))
```

```{r}
#ordering days 
df$day <- ordered(df$day, levels=c("Monday", "Tuesday", "Wednesday", "Thursday", 
"Friday", "Saturday", "Sunday"))
```

```{r hist}
holiday <- traffic %>% filter(holiday != "None")

ggplot(holiday, aes(x = holiday)) +
  geom_bar()

#there are few holidays 
```
```{r hist}
ggplot(df, aes(y = temp)) +
  geom_boxplot()

#there is an outlier we'll have to chop it 
```
```{r hist}
ggplot(traffic, aes(x = weather_main)) +
  geom_bar()
```
```{r hist}
ggplot(traffic, aes(x = weather_description)) +
  geom_bar()
```

```{r hist}
ggplot(traffic, aes(x = clouds_all)) +
  geom_bar()
```
```{r numerical}
nums <- df %>% select_if(is.numeric)
plot(nums)
```
```{r}
plot(df$snow_1h, df$traffic_volume)
```

```{r bp}
#boxplot
ggplot(traffic, aes(x=weather_main, y=traffic_volume, group=weather_main)) + 
  geom_boxplot()
```
```{r}
ggplot(traffic, aes(x = weather_description, y = traffic_volume, group = weather_description)) +
  geom_boxplot()
```
```{r bp}
#boxplot
ggplot(holiday, aes(x=holiday, y=traffic_volume, group=holiday)) + 
  geom_boxplot()
```
```{r}
ggplot(df, aes(x = day, y = traffic_volume, group = day)) +
  geom_boxplot()
```
```{r}
ggplot(df, aes(x = tod, y = traffic_volume, group = tod)) +
  geom_boxplot()
```
```{r}
ggplot(df, aes(x = year, y = traffic_volume)) +
  geom_boxplot() 
```

```{r}
slr <- lm(traffic_volume ~ day, data = df)
summary(slr)

plot(df$day, df$traffic_volume)
abline(lm(traffic_volume ~ day, data = df))
```
```{r}
df$temp = ifelse(df$temp < -50, 0, df$temp)

plot(df$temp, df$traffic_volume)
abline(lm(traffic_volume ~ temp, data = df))
```
```{r}
plot(df$hour2, df$traffic_volume)
abline(lm(traffic_volume ~ hour2, data = df))
```
```{r}
pr <- lm(traffic_volume ~ temp + I(hour2^2) + I(hour2^3), data = df)
summary(pr)

plot(df$hour2, df$traffic_volume)
lines(sort(df$hour2), fitted(pr)[order(df$hour2)], col='red', type='b') 
```


```{r tree}

```

































```{r daily info}
#taking the average of each variable by date
daily_mean_temp <- df %>% group_by(date) %>% summarize(temp = mean(temp))
daily_mean_vol <- traffic %>% group_by(date) %>% summarize(vol = mean(traffic_volume))
daily_mean_rain <- traffic %>% group_by(date) %>% summarize(rain = mean(rain_1h))
daily_mean_snow <- traffic %>% group_by(date) %>% summarize(snow = mean(snow_1h))
daily_mean_cloud <- traffic %>% group_by(date) %>% summarize(cloud = mean(clouds_all))
is_holiday <- df %>% group_by(date) %>% summarize(holiday = mean(holiday))
daily_day <- traffic %>% group_by(date) %>% 
```

```{r weather}
x <- traffic %>% group_by(date) %>% select(weather_main)

daily_weather <- x %>% group_by(date) %>% summarize(weather = calculate_mode(weather_main))
```

```{r average df}
avg <- merge(daily_weather, merge(is_holiday, merge(daily_mean_cloud, merge(daily_mean_snow, merge(daily_mean_rain,merge(daily_mean_temp, daily_mean_vol))))))
```

```{r avg format}
avg %>% 
  mutate(
    holiday = ifelse(holiday != 0, 1, 0),
  )
```
```{r sample}
sample <- sample_n(df, 1000, replace = FALSE)
```