---
title: "final"
author: "esther"
date: "11/17/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(ggplot2)
library(dplyr)
library(lubridate)
library(repr)

options(repr.plot.width=5, repr.plot.height=4)
```

```{r https://stackoverflow.com/a/46846474 source}
#mode of factor variables
calculate_mode <- function(x) {
  uniqx <- unique(x)
  uniqx[which.max(tabulate(match(x, uniqx)))]
}
```

```{r https://stackoverflow.com/a/9501225 source}
getSeason <- function(DATES) {
    WS <- as.Date("2012-12-15", format = "%Y-%m-%d") # Winter Solstice
    SE <- as.Date("2012-3-15",  format = "%Y-%m-%d") # Spring Equinox
    SS <- as.Date("2012-6-15",  format = "%Y-%m-%d") # Summer Solstice
    FE <- as.Date("2012-9-15",  format = "%Y-%m-%d") # Fall Equinox

    # Convert dates from any year to 2012 dates
    d <- as.Date(strftime(DATES, format="2012-%m-%d"))

    ifelse (d >= WS | d < SE, "Winter",
      ifelse (d >= SE & d < SS, "Spring",
        ifelse (d >= SS & d < FE, "Summer", "Fall")))
}
```
The Metro Interstate Traffic volume dataset is from UCI Machine Learning Repostitory, data is taken from Minneapolis-St Paul, MN traffic volume for westbound I-94. There are nine variables and 48,204 observations of every hour from 2012 to 2018. 

Variables:
holiday - categorical indicating if there was a holiday
temperature - temperature in kelvin
rain_1h - rain in mm that occured in the hour
snow_1h snow in mm that occured in the hour
clouds_all - cloud cover percentage
weather_main - description of weather
weather_descrription - long description
date_time - hour of data collected
traffic_volume - numeric hourly average true range of traffic volume

```{r data}
#read in data
traffic <- read_csv("Metro_Interstate_Traffic_Volume.csv")
head(traffic)
dim(traffic)
summary(traffic$traffic_volume)
```

```{r data edit}
unique(traffic$weather_main)
#change columns into new dataframe df
traffic_2 <- traffic %>%
        mutate(
          temp = temp - 270.15,
          weather_main = factor(weather_main),
          holiday = ifelse(holiday == "None", 0, holiday) %>% as.numeric,
          holiday = ifelse(is.na(holiday), 1, 0) %>% as.numeric)
```
```{r adding columns}
#new columns
traffic_2$year <- strftime(traffic_2$date_time, format="%Y")

traffic_2$season <- getSeason(traffic_2$date_time)

traffic_2$month <- month(traffic_2$date_time)

traffic_2$date <- as.Date(traffic_2$date_time, "%m/%d/%Y")

traffic_2$weekday <- weekdays(traffic_2$date)

traffic_2$weekday <- ordered(traffic_2$date, levels=c("Monday", "Tuesday", "Wednesday", "Thursday", 
"Friday", "Saturday", "Sunday"))

# traffic_2$tod <- ifelse(traffic_2$hournum >= 03 & traffic_2$hournum <= 07 , "Early Morning",
#             ifelse(traffic_2$hournum > 07 & traffic_2$hournum <= 11, "Morning",
#             ifelse(traffic_2$hournum > 11 & traffic_2$hournum <= 17, "Afternoon", 
#             ifelse(traffic_2$hournum > 17 & traffic_2$hournum <= 20, "Evening", "Night"))))
# 
# traffic_2$tod <- ordered(traffic_2$tod, levels = c("Early Morning", "Morning", "Afternoon", "Evening", "Night"))
# 
# traffic_2 %>% mutate(tod = as.factor(tod))

traffic_2$hour <- strftime(traffic_2$date_time, format="%H", tz = "CST")

traffic_2$temp = ifelse(traffic_2$temp < -50, 0, traffic_2$temp)
```



```{r histograms}
ggplot(traffic, aes(x = traffic_volume)) +
  geom_histogram(bins = 40)

holiday <- traffic %>% filter(holiday != "None")

ggplot(holiday, aes(x = holiday)) +
  geom_bar()

ggplot(traffic, aes(x = weather_main)) +
  geom_bar() 

ggplot(traffic, aes(x = weather_description)) +
  geom_bar()

ggplot(traffic, aes(x = clouds_all)) +
  geom_bar()
```


```{r numerical plots}
plot(traffic$clouds_all, traffic$traffic_volume)
plot(traffic$snow_1h, traffic$traffic_volume)
plot(traffic$rain_1h, traffic$traffic_volume)
```

```{r weather and volume}
ggplot(traffic, aes(x=weather_main, y=traffic_volume, group=weather_main)) + 
  geom_boxplot()

ggplot(traffic, aes(x = weather_main, y = traffic_volume, group = weather_main)) +
  geom_bar(stat = "identity")

traffic %>% filter(weather_main == "Clouds") %>% summarise(mean = mean(traffic_volume))
traffic %>% filter(weather_main == "Drizzle") %>% summarise(mean = mean(traffic_volume))
```

```{r weather desc and volume}
ggplot(traffic, aes(x = weather_description, y = traffic_volume, group = weather_description)) +
  geom_boxplot()

ggplot(holiday, aes(x=holiday, y=traffic_volume, group=holiday)) + 
  geom_boxplot()

ggplot(holiday, aes(x = holiday, y = traffic_volume, group = holiday)) +
  geom_bar(stat = "identity")

ggplot(traffic_2, aes(x = weekday, y = traffic_volume)) +
  geom_boxplot()

ggplot(traffic_2, aes(x = weekday, y = traffic_volume, group = weekday)) +
  geom_bar(stat = "identity")

# ggplot(traffic, aes(x = tod, y = traffic_volume, group = tod)) +
#   geom_boxplot()
# 
# ggplot(traffic, aes(x = tod, y = traffic_volume, group = tod)) +
#   geom_bar(stat = "identity")

ggplot(traffic_2, aes(x = year, y = traffic_volume)) +
  geom_boxplot()

ggplot(traffic_2, aes(x = month, y = traffic_volume, group = month)) +
  geom_boxplot()

ggplot(traffic_2, aes(x = hour, y = traffic_volume, group = hour)) +
  geom_boxplot()
```


```{r}
plot(traffic$temp, traffic$traffic_volume)

```
```{r}
plot(traffic$clouds_all, traffic$traffic_volume)
plot(traffic$weather_main, traffic$temp)
```

```{r data into years}
# twelve <- df %>% filter(year == 2012)
# thirteen <- df %>% filter(year == 2013)
# fourteen <- df %>% filter(year == 2014)
# fifteen <- df %>% filter(year == 2015)
# sixteen <- df %>% filter(year == 2016)
# seventeen <- df %>% filter(year == 2017)
# eighteen <- df %>% filter(year == 2018)
```


```{r}
ggplot(traffic_2, aes(x = hour, y = traffic_volume)) +
  geom_point() +
    facet_wrap(~day)
```
```{r}
traffic_2_select <- traffic_2 %>% filter(year ==2018)

ggplot(traffic_2_select, aes(x = date, y = traffic_volume)) +
  geom_line()

```
```{r}
ggplot(traffic, aes(x = weather_main)) +
  geom_bar()


time_plot <- ggplot(traffic_2_select, aes(x = date, y = traffic_volume)) +
	geom_line() +
	scale_x_date(date_labels = "%Y", date_breaks = "1 year") +
	theme_classic()

time_plot

decomp_2 <- ggplot(traffic_2, aes(x = date, y = traffic_volume)) +
	geom_line() +
	geom_smooth(method = "loess", se = FALSE, span = 0.6) +
	theme_classic()

decomp_2
```


